\textcolor{Navy}{\textsc{Convexity adjustments with a bit of Malliavin}}

### 2022-10-06 Latex/paper.tex
# first steps cms
# cms convexity

\hspace{1cm}
\fboxsep1em
\fbox{\parbox{435pt}
{
\begin{equation}\label{ca_approximation_futures}
CA(t,t_0,t_1) \approx \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) \frac{P_{ois}(0,t_2)}{P_{ois}(0,t_0)} - G(t_0,t_1) \frac{P_{ois}(0,t_1)}{P_{ois}(0,t_0)}\right) \int_{0}^{t_0}  \beta(s,t_0, x_0, \hat{y}_s) \nu(s,t_2) ds. 
\end{equation}
}}
\begin{example}\label{example_ca_future}
Let us to set 
\begin{align*}
g(T) &= \exp(-kT) \\
h(t) &= \sigma
\end{align*}
\end{example}
With this parametrization, the Cheyette model is equiavalent to Hull-White model. It is easy to show that under this parametrization, the convexity adjustment (\ref{ca_approximation_futures}) is
\begin{align*}
CA(t_0,t_1) & \approx \frac{\sigma^{2} \exp(-k t_0)  P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(\frac{1 - \exp(- k t_0)}{k^{2}} - \frac{t_0 \exp(-k t_2)}{k} \right).   
\end{align*}
In the next figure, we can check the accuracy of the last formula versus montecarlo. The parameters that we have used are $\sigma=0.05$, $k=0.003$ and flat curve with level $r=0.01$.

@@

\subsection{OIS futures}
In this we will derive the covexity adjustment for overnight-indexed swap (OIS) under. We will define for $t_0 < t_1$
\begin{align*}
R(t_0,t_1) &= \frac{\exp\left(\int_{t_0}^{t_1}r_{ois,u} du \right) - 1}{t_1 - t_0} \\
R_{avg}(t_0,t_1) &= \frac{\int_{t_0}^{t_1}r_{ois,u} du}{t_1 - t_0}.
\end{align*}
We must observe, that both $R(\cdot,t_0,t_1)$ and $R_{avg}(\cdot,t_0,t_1)$  are not predictible and they are only observable in $t_1$. However, since
$R(\cdot,t_0,t_1)$ and $R_{avg}(\cdot,t_0,t_1)$ are flows that will be payed in $t_1$, we can considerer the expected value under the measure $\mathbb{Q}^{t_1}$ that will observable during whole period $[t_0, t_1]$. Let us define the next $\mathbb{Q}^{t_1}$ martingales  
\begin{align*} 
\bar{R}(t,t_0,t_1) &= \mathbb{E}^{\mathbb{Q}}\left( R(t_0,t_1)  \right) \\
\bar{R}_{avg}(t,t_0,t_1) &= \mathbb{E}^{\mathbb{Q}}\left( R_{avg}(t_0,t_1)  \right).
\end{align*}
We must do several obeservations. The first observation is that
\begin{align*}
F(t,t_0,t_1)=\mathbb{E}_{t}^{\mathbb{Q}}\left(\exp\left(-\int_{t}^{t_0} r_{ois,u} du \right) R(t_0,t_1) \right) &= \frac{\frac{P_{ois}(t,t_0)}{P_{ois}(t,t_1)} - 1}{t_1 - t_0}, \quad t \in [0,t_0] \\
F(t,t_0,t_1)=\mathbb{E}_{t}^{\mathbb{Q}}\left(\exp\left(-\int_{t}^{t_0} r_{ois,u} du \right) R(t_0,t_1) \right) &= \frac{1}{t_1 - t_0} \left(\frac{\left( \exp\left(\int_{t}^{t_1}r_{ois,u} du\right)\right)}{P_{ois}(t_0,t)}-1\right), \quad  t \in  [t_0, t_1].
\end{align*}
The can define the convexity adjustment for $ R(t_0,t_1)$ the next way
\begin{equation}\label{R_ois_ca}
CA_{ois}(t,t_0,t_1) = F(t,t_0,t_1) - \bar{R}(t,t_0,t_1) 
\end{equation}
The second observation, is that under the dyam we must do is that
\begin{equation}\label{R_ois_avg}
\mathbb{E}_{t}^{\mathbb{Q}}\left(R_{avg}(t_0,t_1) \right) = \mathbb{E}_t^{\mathbb{Q}}\left( \frac{\log\left(1+ \delta_{t_0,t_1} R(t_0,t_1) \right)}{\delta_{t_0,t_1}} \right)  
\end{equation}
In order to compute $\mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1\right)$, we will define
\begin{equation*}
I(t_0,t_1) = \int_{t_0}^{t_1} r_s ds
\end{equation*}
then for $t < t_0$ from (\ref{clark-okone}) and (\ref{malliavin_derive_x}), we have the next representation for $I(t_0,t_1)$
\begin{align}\label{apprx_I_t0_t1}
I(t_0,t_1) &= \mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right) + \int_{0}^{t_1}\int_{\max(s, t_{0})}^{t_1}  \mathbb{E}_s^{\mathbb{Q}}\left(\beta(s,u,x_s,y_s) M(s,u) \right) du dW_s^{\mathbb{Q}} \nonumber \\
&\approx \mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right) + \int_{0}^{t_1}\int_{\max(s, t_{0})}^{t_1} \beta(s,u,x_0,\hat{y}_s) du dW_s^{\mathbb{Q}}\nonumber \\
&= \mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right) + \int_{0}^{t_1} g(s)h(s,x_0,\hat{y}_s)\int_{\max(s, t_{0})}^{t_1} \exp\left( -\int_{s}^{u} k_{s^{\prime}} ds^{\prime}\right) du dW_s^{\mathbb{Q}}.
\end{align}
Therefore, we can get the next approximation
\begin{align*}
1 + \delta_{t_0,t_1}\mathbb{E}^{\mathbb{Q}^{t_1}}\left(R(t_0,t_1)\right) &= \mathbb{E}^{\mathbb{Q}^{t_1}}\left( \exp(I(t_0,t_1)) \right)\\
&= \exp\left(\mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right)\right)  \mathbb{E}^{\mathbb{Q}^{t_1}}\left(\exp\left(\int_{0}^{t_1} \Gamma(s,t_0,t_1)dW_s^{\mathbb{Q}}\right)\right)
\end{align*}
where $\Gamma(s,t_0,t_1)= g(s)h(s,x_0,\hat{y}_s)\int_{\max(s, t_{0})}^{t_1} \exp\left( -\int_{s}^{u} k_{s^{\prime}} ds^{\prime}\right)du$. Now, from the Girsanov's theorem we get that
\begin{equation*}
1 + \delta_{t_0,t_1}\mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1)\right) \approx \exp\left(\mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right)\right)\exp\left(\int_{0}^{t_1}\frac{\Gamma^{2}(s,t_0,t_1)}{2} ds\right)
\end{equation*}
Therefore we have that

@@

\subsection{CMSs}

>>>>

\hspace{1cm}
\fboxsep1em
\fbox{\parbox{435pt}
{
\begin{equation}\label{ca_approximation_futures}
CA(t,t_0,t_1) \approx \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) \frac{P_{ois}(0,t_2)}{P_{ois}(0,t_0)} - G(t_0,t_1) \frac{P_{ois}(0,t_1)}{P_{ois}(0,t_0)}\right) \int_{0}^{t_0}  \beta(s,t_0, x_0, \hat{y}_s) \nu(s,t_2) ds. 
\end{equation}
}}
\begin{example}\label{example_ca_future}
Let us to set 
\begin{align*}
g(T) &= \exp(-kT) \\
h(t) &= \sigma
\end{align*}
\end{example}
With this parametrization, the Cheyette model is equiavalent to Hull-White model. It is easy to show that under this parametrization, the convexity adjustment (\ref{ca_approximation_futures}) is
\begin{align*}
CA(t_0,t_1) & \approx \frac{\sigma^{2} \exp(-k t_0)  P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(\frac{1 - \exp(- k t_0)}{k^{2}} - \frac{t_0 \exp(-k t_2)}{k} \right).   
\end{align*}
In the next figure, we can check the accuracy of the last formula versus montecarlo. The parameters that we have used are $\sigma=0.015$, $k=0.003$ and flat curve with level $r=0.01$.

@@

\subsection{OIS futures}
In this we will derive the covexity adjustment for overnight-indexed swap (OIS) under. We will define for $t_0 < t_1$
\begin{align*}
R(t_0,t_1) &= \frac{\exp\left(\int_{t_0}^{t_1}r_{ois,u} du \right) - 1}{t_1 - t_0} \\
R_{avg}(t_0,t_1) &= \frac{\int_{t_0}^{t_1}r_{ois,u} du}{t_1 - t_0}.
\end{align*}
We must observe, that both $R(\cdot,t_0,t_1)$ and $R_{avg}(\cdot,t_0,t_1)$  are not predictible and they are only observable in $t_1$. However, since
$R(\cdot,t_0,t_1)$ and $R_{avg}(\cdot,t_0,t_1)$ are flows that will be payed in $t_1$, we can considerer the expected value under the measure $\mathbb{Q}^{t_1}$ that will observable during whole period $[t_0, t_1]$. Let us define the next $\mathbb{Q}$ martingales  
\begin{align*} 
\bar{R}(t,t_0,t_1) &= \mathbb{E}^{\mathbb{Q}}\left( R(t_0,t_1)  \right) \\
\bar{R}_{avg}(t,t_0,t_1) &= \mathbb{E}^{\mathbb{Q}}\left( R_{avg}(t_0,t_1)  \right).
\end{align*}
We must do several obeservations. The first observation is that
\begin{align*}
F(t,t_0,t_1)=\mathbb{E}_{t}^{\mathbb{Q}}\left(\exp\left(-\int_{t}^{t_0} r_{ois,u} du \right) R(t_0,t_1) \right) &= \frac{\frac{P_{ois}(t,t_0)}{P_{ois}(t,t_1)} - 1}{t_1 - t_0}, \quad t \in [0,t_0] \\
F(t,t_0,t_1)=\mathbb{E}_{t}^{\mathbb{Q}}\left(\exp\left(-\int_{t}^{t_0} r_{ois,u} du \right) R(t_0,t_1) \right) &= \frac{1}{t_1 - t_0} \left(\frac{\left( \exp\left(\int_{t}^{t_1}r_{ois,u} du\right)\right)}{P_{ois}(t_0,t)}-1\right), \quad  t \in  [t_0, t_1].
\end{align*}
The can define the convexity adjustment for $ R(t_0,t_1)$ the next way
\begin{equation}\label{R_ois_ca}
CA_{ois}(t,t_0,t_1) = F(t,t_0,t_1) - \bar{R}(t,t_0,t_1) 
\end{equation}
The second observation, is that under the dyam we must do is that
\begin{equation}\label{R_ois_avg}
\mathbb{E}_{t}^{\mathbb{Q}}\left(R_{avg}(t_0,t_1) \right) = \mathbb{E}_t^{\mathbb{Q}}\left( \frac{\log\left(1+ \delta_{t_0,t_1} R(t_0,t_1) \right)}{\delta_{t_0,t_1}} \right)  
\end{equation}
In order to compute $\mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1\right)$, we will define
\begin{equation*}
I(t_0,t_1) = \int_{t_0}^{t_1} r_s ds
\end{equation*}
then for $t < t_0$ from (\ref{clark-okone}) and (\ref{malliavin_derive_x}), we have the next representation for $I(t_0,t_1)$
\begin{align}\label{apprx_I_t0_t1}
I(t_0,t_1) &= \mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right) + \int_{0}^{t_1}\int_{\max(s, t_{0})}^{t_1}  \mathbb{E}_s^{\mathbb{Q}}\left(\beta(s,u,x_s,y_s) M(s,u) \right) du dW_s^{\mathbb{Q}} \nonumber \\
&\approx \mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right) + \int_{0}^{t_1}\int_{\max(s, t_{0})}^{t_1} \beta(s,u,x_0,\hat{y}_s) du dW_s^{\mathbb{Q}}\nonumber \\
&= \mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right) + \int_{0}^{t_1} g(s)h(s,x_0,\hat{y}_s)\int_{\max(s, t_{0})}^{t_1} \exp\left( -\int_{s}^{u} k_{s^{\prime}} ds^{\prime}\right) du dW_s^{\mathbb{Q}}.
\end{align}
Therefore, we can get the next approximation
\begin{align*}
1 + \delta_{t_0,t_1}\mathbb{E}^{\mathbb{Q}^{t_1}}\left(R(t_0,t_1)\right) &= \mathbb{E}^{\mathbb{Q}^{t_1}}\left( \exp(I(t_0,t_1)) \right)\\
&= \exp\left(\mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right)\right)  \mathbb{E}^{\mathbb{Q}^{t_1}}\left(\exp\left(\int_{0}^{t_1} \Gamma(s,t_0,t_1)dW_s^{\mathbb{Q}}\right)\right)
\end{align*}
where $\Gamma(s,t_0,t_1)= g(s)h(s,x_0,\hat{y}_s)\int_{\max(s, t_{0})}^{t_1} \exp\left( -\int_{s}^{u} k_{s^{\prime}} ds^{\prime}\right)du$. Now, from the Girsanov's theorem we get that
\begin{equation*}
1 + \delta_{t_0,t_1}\mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1)\right) \approx \exp\left(\mathbb{E}^{\mathbb{Q}}\left(I(t_0,t_1)\right)\right)\exp\left(\int_{0}^{t_1}\frac{\Gamma^{2}(s,t_0,t_1)}{2} ds\right)
\end{equation*}
Therefore we have that

@@

\subsection{CMSs}
We will define the swap rate from $T_a$ to $T_b$ at time $t$
\begin{equation*}
S_{a,b} = \frac{\sum_{i=1}^{n_E}\delta_{t^{E}_{i-1}, t^{E}_i} L^{E}(t,t^{E}_{i-1}, t^{E}_{i}) P_{ois}(t,t^{E}_{i})}{01(t,T_a,T_b)}
\end{equation*}
where
\begin{align*}
01(t,T_a,T_b) = \sum_{j=1}^{n_f} \delta_{t^{f}_{i-1}, t^{f}_i} P_{ois}(t,t^{f}_{j}) \\
T_a=T^{E}_0 < T^{E}_i< \cdots < T^{E}_{n_E}=T_b \quad i=0,\cdots,n_E&  \\
T_a=T^{f}_0 < T^{f}_j< \cdots < T^{f}_{n_f}=T_b \quad j=0,\cdots,n_f&
\end{align*}
We suppose, that we have a flow in $T_a < T_p < T_b$ with value $S_{a,b}(T_a)$. Then, the present value of this flow
\begin{equation*}
P_{ois}(0,T_a)\mathbb{E}^{\mathbb{Q}^{T_{a}}}\left(S_{a,b}(T_a) P_{ois}(T_a,T_p) \right) = 01(0,T_a,T_b) \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) \frac{P_{ois}(T_a,T_p)}{01(T_a,T_a,T_b) }\right) 
\end{equation*}
therefore, we have that
\begin{align}\label{cms_expedted_value}
\mathbb{E}^{\mathbb{Q}^{T_{a}}}\left(S_{a,b}(T_a) P_{ois}(T_a,T_p) \right) &=  \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) \frac{M(T_a)}{M(0)}\right) \nonumber \\
& = S_{a,b}(0) + \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) \left(\frac{M(0, T_p)}{M(T_a, T_p)} - 1\right) \right)  
\end{align} 
where $M(t)= \frac{01(t,T_a,T_b)}{P_{ois}(t,T_p)} $. From (\ref{cms_expedted_value}), we will define the convexity adjusment for a CMS the next way
\begin{equation}\label{cms_convexity}
CA_{CMS}(T_a,T_b) = \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) \left(\frac{M(0, T_p)}{M(T_a, T_p)} - 1\right) \right). 
\end{equation}
In order to reduce the complexity to explain the movement od the forward curve, it is a common pratice to asumme that
\begin{equation*}
M(t) = f(S_{a,b}(t))
\end{equation*}
where $f(\cdot)$ is known as mapping function. There is a extensive literature how to define this function (see Piterbarg Volume III). Then under this assumption and applying Clark-Ocone to $f(S_{a,b}(t))$ we get
\begin{align*}
f(S_{a,b}(T_a)) &= \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a)\right) + \int_{0}^{T_a} \mathbb{E}_s^{01(a,b)}\left( D_s (f(S_{a,b}(T_a))) \right) dW^{01(a,b)}_{s} \\
&= 1 + \int_{0}^{T_a} \mathbb{E}_s^{01(a,b)}\left(f^{\prime}(S_{a,b}(T_a)) \partial_{x}S_{a,b}(T_a) D_s x_{T_a} \right) dW^{01(a,b)}_{s}. 
\end{align*}
Therefore, we have that
\begin{equation*}
CA_{CMS}(T_a,T_b) = \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) \int_{0}^{T_a} \mathbb{E}_s^{01(a,b)}\left( f^{\prime}(S_{a,b}(T_a)) \partial_{x}S_{a,b}(T_a) D_s x_{T_a} \right) dW^{01(a,b)}_{s} \right).
\end{equation*}
Now if we apply the adjoint of the Malliavin derivative (see Elisa-David) we get

\begin{align}
CA_{CMS}(T_a,T_b) &= \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left( \int_{0}^{T_a} \partial_{x}S_{a,b}(T_a) D_s x_{T_a} \mathbb{E}_s^{01(a,b)}\left(f^{\prime}(S_{a,b}(T_a)) \partial_{x}S_{a,b}(T_a) D_s x_{T_a} \right) ds \right) \nonumber \\
&\approx  \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left( \int_{0}^{T_a} \left(\partial_{x}S_{a,b}(0)\right)^{2} f^{\prime}(S_{a,b}(0))  D_s x_{T_a} \mathbb{E}_s^{01(a,b)}\left(D_s x_{T_a} \right) ds \right)
\end{align}
